The Decline of Financial Wellness

### 2023-03-09 outputs/paper/paper3.qmd
draft data section complete

---
title: "The Decline of Financial Wellness"
title-block-banner: true
abstract: "TBD"
thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/INF3104-3-US-Survey.git"
author: 
  - Michaela Drouillard
  - Christina Wei
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: ref.bib
---

@@

-   Simulated data - Christina

-   Data validation tests - Christina

@@

-   financial change

-   financial compare

-   Supplementary survey

@@

The super upper class are in a bubble and delusional

@@

Data was cleaned and analyzed using the open source statistically programming language R [@r], using functionalities from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra] and `knitr` [@rKnitr]. Details of the data extraction and cleaning processes are discussed in the subsections below.

## General Social Survey

Since 1972, the General Social Survey (GSS) has been tracking trends in public opinion through in-person interviews. Due to the COVID-19 pandemic, the GSS had to change their data collection from in-person to address-based sampling and a push-to-web methodology, meaning that most of the interview was conducted with an online self-administered questionnaire. When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey. T

### Surveys Related to Financial Attitudes

### Respondent Demographics

@@

## Social Class

@@

## Financial Compare

@@

##

@@

## Gaps

## Translations

## Affordances: New Experiments

## Impact on our study

>>>>

---
title: "The Decline of Financial Wellness"
title-block-banner: true
abstract: "TBD"
thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/INF3104-3-US-Survey.git"
author: 
  - Christina Wei
  - Michaela Drouillard
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: ref.bib
---

@@

-   Simulated data

-   Data validation tests

-   README

@@

-   Supplementary survey: The survey should have an introductory section and include the details of a contact person. The survey questions should be well constructed and appropriate to the task. The questions should have an appropriate ordering. A final section should thank the respondent

-   Introduction

-   Discussions

-   Limitations & Future Work

To Do

-   Abstract

-   Conclusion

-   Title

@@

The upper class are in a bubble and delusional

@@

Data used in this pare are retrieved from the US General Social Survey (GSS) from NORC at the University of Chicago [@gss]. We retrieved demographic data as well as survey questions related to financial wellness from 1972 to 2021.

## Source Data

Specifically, we retrieved the following demographic data (TODO: explain these values)

-   cohort

-   degree

-   class

-   class1

-   sexbirth1

-   martial

-   realinc

-   income16

For survey questions related to financial wellness, we retrieved the following data points (TODO: explain these values):

-   confinan

-   satfin

-   finalter

-   finrela

## Data Limitations

Since 1972, the General Social Survey (GSS) has been tracking trends in public opinion through in-person interviews. Due to the COVID-19 pandemic, the GSS had to change their data collection from in-person to address-based sampling and a push-to-web methodology, meaning that most of the interview was conducted with an online self-administered questionnaire. When using this primarily mail-based communication, the 2021 GSS asks for "the person with the most recent birthday, rather than a random person in the household" to respond to the survey.

### Gaps

### Translations

### Affordances: New Experiments

### Impact on our study

## Data Cleaning

Data was cleaned and analyzed using the open source statistically programming language R [@r], using functionalities from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `haven` [@rHaven], `here` [@rHere], `formattable` [@rFormattable], `kableExtra` [@rKableExtra] and `knitr` [@rKnitr].

After downloading and filtering for the desired source data values from GSS[^1], we performed data cleaning on each one of the columns based on value definitions as defined in GSS codebook [@gssCodebook], as well as cleaning up column names. For example, the column `sexbirth1` has been renamed to `gender`, with the value 1 updated to "Male", and the value 2 updated to "Female" based on mapping in the codebook.

[^1]: https://gss.norc.org/documents/stata/GSS_stata.zip

Overall, there are 68,846 observations collected across 33 surveys administered over the years. On average there are 2086 observations per survey, with the maximum number of 4510 observations in year 2006, and minimum number of 1372 observations in year 1990. The standard deviation for the number of observations per survey is 788 (see @tbl-statistics).

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-statistics
#| tbl-cap: Summary statistics for number of observations in surveys
combined_data |>
  group_by(year) |>
  count() |>
  ungroup() |>
  summarise(
    mean = mean(n),
    max = max(n),
    min = min(n),
    sd = sd(n)
  ) |>
  kable(
    booktabs = TRUE,
    digits = 1
  )
```

@@

## Social Class (still need to fix up)

@@

## Financial Compare (old)

@@

@@

Additional financial variables in survey for analysis

Connect with US census data to check representativeness

Connect with other macroeconomic indicators - e.g. inflation, stock markets

Collect more data to supplement our findings -\> link to survey <https://docs.google.com/forms/d/e/1FAIpQLSdqe87Bet1PyaCezPxN_t7A1smQCpSap6EYjG2LrTN0K_CX9A/viewform>
